The Decreasing Orbital Period of WASP-4\lowercase{b}

### 2018-12-06 paper/ms.tex
update tex w/ dynamics limits

Another mechanism to excite the hot Jupiter's eccentricity is the
Kozai-Lidov mechanism \citep{lidov_evolution_1962,kozai_secular_1962}.
In this case, the orbital plane of the outer companion, ``c'', would
need to be inclined relative to that of the hot Jupiter, ``b'', by at
least $\sin^{-1} \sqrt{2/5} \approx 39^\circ$.
\citet{bailey_understanding_2019} point out how the joint requirements
of Kozai-Lidov oscillations with a non-detection of an additional RV
signal can be combined to produce lower and upper limits on the
companion's mass.  For the Kozai-Lidov mechanism to operate, from
\citet{bailey_understanding_2019} Equation 20 we need
\begin{equation}
  M_{\rm c} > 7.7\,M_\oplus
  \times \left( \frac{a_{\rm c}}{a_{\rm b}} \right)^{3/2},
  \label{eq:kozai_bound}
\end{equation}
where $M_{\rm b,c}$ and $a_{\rm b,c}$ are the mass and semi-major axis
of WASP-4b and the hypothetical WASP-4c, and we have assumed $k_{2,{\rm
b}}\approx 0.6$, for $k_{2,{\rm b}}$ WASP-4b's Love number.  For the RV
signal of the companion to remain undetected, it would need to be in the
residual $(O-C)_{\rm RV} = 15.2\,{\rm m s}^{-1}$ reported by
\citet{triaud_spin-orbit_2010}.  Again following
\citet{bailey_understanding_2019}, this implies
\begin{align}
  M_{\rm c} &<
  (O-C)_{\rm RV}
  \left( \frac{M_\star a_b}{G} \right)^{1/2}
  \left( \frac{a_c}{a_b} \right)^{1/2}
  f^{-1/2}
  \nonumber
  \\
  &< 
  24.2\,M_\oplus
  \times 
  \left( \frac{a_c}{a_b} \right)^{1/2}
  f^{-1/2},
  \label{eq:rv_bound}
\end{align}
for $f(e_{\rm c},\omega_{\rm c}, i_{\rm c}) \propto sin^2 i_{\rm c}$ a
geometric prefactor that depends on the argument of periastron and
inclination of the exterior companion
(\citealt{bailey_understanding_2019} Equation 23).  Since WASP-4b
transits along the line of sight, $f$ can be arbitrarily small, and both
of the preceding limits can be easily satisfied.  While the mechanism
could produce the needed eccentricity, it would be more satifying if
long-term radial velocity monitoring showed an additional planet in the
system.  The chances of one existing with a mass between 1-13\,$M_{\rm
Jup}$ and orbital semi-major axis between 1-20\,AU is about 50\%, and a
linear trend is currently present at $\sim 2\sigma$
\citep{knutson_friends_2014}.

\paragraph{Fluctuations in the gravitational potential from convection}
A separate mechanism that might produce the eccentricity is simply
stellar convection \citep[][Section 7]{phinney_pulsars_1992}.  Phinney
presents this idea by analogy with a pendulum, which is never really
at rest. Instead, collisions with air molecules randomly perturb its
momentum and energy, and the equipartition theorem of statistical
mechanics tells us the average energy is $\langle E_{\rm pend} \rangle
= k T_{\rm air}$.  In hot Jupiter systems, the ``air molecules'' are
convective eddies, and their associated kinetic energy leads the orbit
to never have exactly zero eccentricity.  In particular, from
\citealt{phinney_pulsars_1992} Equation (7.33),
\begin{equation}
  \langle e^2 \rangle =
  \frac{ 2 \langle E_{\rm e} \rangle }{\mu n^2 a^2}
  = 6.8\times10^{-5}
  \frac{(L^2 R_{\rm conv}^2 M_{\rm conv}^2)^{1/3}}{\mu n^2 a^2},
\end{equation}
for $L$ the stellar luminosity, $R_{\rm conv}$ ($M_{\rm conv}$) the
width (mass) of the convective region,  $\mu$ the reduced mass, $n$
the orbital frequency, and $a$ the semi-major axis.  For WASP-4,
convection in the star is more important than convection in the
planet, and $\langle e^2 \rangle^{1/2} < 10^{-5}$.

>>>>

Another mechanism to excite the hot Jupiter's eccentricity is the
Kozai-Lidov mechanism \citep{lidov_evolution_1962,kozai_secular_1962}.
In this case, the orbital plane of the outer companion, ``c'', would
need to be inclined relative to that of the hot Jupiter, ``b'', by at
least $\sin^{-1} \sqrt{2/5} \approx 39^\circ$.
\citet{bailey_understanding_2019} point out how the joint requirements
of Kozai-Lidov oscillations with a non-detection of an additional RV
signal can be combined to produce lower and upper limits on the
companion's mass.  For the Kozai-Lidov mechanism to operate, from
\citet{bailey_understanding_2019} Equation 20 we need
\begin{equation}
  M_{\rm c} > 7.7\,M_\oplus
  \times \left( \frac{a_{\rm c}}{a_{\rm b}} \right)^{3/2},
  \label{eq:kozai_bound}
\end{equation}
where $M_{\rm b,c}$ and $a_{\rm b,c}$ are the mass and semi-major axis
of WASP-4b and the hypothetical WASP-4c, and we have assumed WASP-4b's
Love number is $k_{2,{\rm b}}\approx 0.6$.  For the RV signal of the
companion to remain undetected, it would need to be in the residual
$({\rm O-C})_{\rm RV} = 15.2\,{\rm m s}^{-1}$ reported by
\citet{triaud_spin-orbit_2010}.  Again following
\citet{bailey_understanding_2019}, this implies
\begin{align}
  M_{\rm c} &<
  ({\rm O-C})_{\rm RV}
  \left( \frac{M_\star a_b}{G} \right)^{1/2}
  \left( \frac{a_c}{a_b} \right)^{1/2}
  f^{-1/2}
  \nonumber
  \\
  M_{\rm c} &< 
  24.2\,M_\oplus
  \times 
  \left( \frac{a_c}{a_b} \right)^{1/2}
  f^{-1/2},
  \label{eq:rv_bound}
\end{align}
for $f(e_{\rm c},\omega_{\rm c}, i_{\rm c}) \propto \sin^2 i_{\rm c}$
a geometric prefactor that depends on the argument of periastron
$\omega_{\rm c}$ and inclination $i_{\rm c}$ of the exterior companion
(\citealt{bailey_understanding_2019} Equation 23).  Since WASP-4b
transits along the line of sight, $f$ can be arbitrarily small, and
both of the preceding limits can be easily satisfied.  This
mechanism could produce the needed eccentricity. However it would be more
satifying if long-term radial velocity monitoring showed an additional
planet in the system.  The chances of one existing with a mass between
1-13\,$M_{\rm Jup}$ and orbital semi-major axis between 1-20\,AU is
about 50\%, and a linear trend is currently present at $\sim 2\sigma$
\citep{knutson_friends_2014}.

\paragraph{Fluctuations in the gravitational potential from convection}
A separate mechanism that might produce the eccentricity is simply
stellar convection \citep[][Section 7]{phinney_pulsars_1992}.  Phinney
presents this idea by analogy with a pendulum, which is never really
at rest. Instead, collisions with air molecules randomly perturb its
momentum and energy, and the equipartition theorem of statistical
mechanics tells us the average energy is $\langle E_{\rm pend} \rangle
= k T_{\rm air}$.  In hot Jupiter systems, the ``air molecules'' are
convective eddies, and their associated kinetic energy leads the orbit
to never have exactly zero eccentricity.  In particular, from
\citealt{phinney_pulsars_1992} Equation 7.33,
\begin{equation}
  \langle e^2 \rangle =
  \frac{ 2 \langle E_{\rm e} \rangle }{\mu n^2 a^2}
  = 6.8\times10^{-5}
  \frac{(L^2 R_{\rm conv}^2 M_{\rm conv}^2)^{1/3}}{\mu n^2 a^2},
\end{equation}
for $L$ the stellar luminosity, $R_{\rm conv}$ ($M_{\rm conv}$) the
width (mass) of the convective region,  $\mu$ the reduced mass, $n$
the orbital frequency, and $a$ the semi-major axis.  For WASP-4,
convection in the star is more important than convection in the
planet, and $\langle e^2 \rangle^{1/2} < 10^{-5}$.
